---
title: "Final Project Milestone 2: Upated Proposal and Analyses"
author: Adam Aleksic
output: pdf_document
---

## Data Set

Data set: https://catalog.data.gov/dataset/racial-and-social-equity-composite-index-a44fc

Understanding the variables: https://data-seattlecitygis.opendata.arcgis.com/datasets/225a4c2c50e94f2cb548a046217f49f7_0?geometry=-122.509%2C47.574%2C-122.164%2C47.655

This data set examines linguistic, racial, ethnic, income, education, and health statistics for census tracts in Seattle.

I want to examine how the percent of English language learners in census tracts correlates with factors like obesity, poverty, education level, asthma, diabetes, and mental health.

## New Research Question

Will census tracts with higher percentages of English language learners have health, income, and education disadvantages relative to census tracts with lower percentages of English language learners? In this study, I plan to use public information released by the city of Seattle to visualize the correlations between ELL status and certain health, education, and income variables. I hypothesize that, because ELL speakers tend to live in institutionally disadvantaged areas and tend to be of institutionally disadvantaged demographics, census tracts with more ELL speakers will average higher incidences of asthma, obesity, diabetes, and mental health issues, and lower education levels and income statuses.

## Analyses

```{r}

library(tidyverse)
library(ggplot2)

seattle <- read.csv("~/Downloads/Gov/Racial_and_Social_Equity_Composite_Index.csv") %>%
 na.omit()

# setting up my data. It's very clean already so I thankfully don't have to do much

```


```{r, education}

education <- seattle %>%
ggplot(aes(x = PCT_ENGLISH_LESSTHAN_VERY_WELL, y = PCT_LESS_BACHELOR_DEGREE)) +
 geom_point(color = "steelblue2") +
 labs(title = "Lack of English Proficiency versus Lack of Education",
      x = "Percentage of English Language Learners",
      y = "Percent with Less Than a Bachelor's Degree") +
 theme_minimal() +
 geom_smooth(method = lm, se = FALSE)
education

# pretty standard scatterplot. I'm doing the same thing for all the other variables so I won't comment much.

education_fit <- lm(PCT_ENGLISH_LESSTHAN_VERY_WELL ~ PCT_LESS_BACHELOR_DEGREE, data = seattle)
education_fit_sum <- summary(education_fit)
education_fit_sum

# finding regression info

```


```{r, obese}

obese <- seattle %>%
ggplot(aes(x = PCT_ENGLISH_LESSTHAN_VERY_WELL, y = PCT_ADULT_OBESE)) +
 geom_point(color = "royalblue4") +
 labs(title = "Lack of English Proficiency versus Obesity",
      x = "Percentage of English Language Learners",
      y = "Percentage of People who are Obese") +
 theme_minimal() +
 geom_smooth(method = lm, se = FALSE)
obese
obese_fit <- lm(PCT_ADULT_OBESE ~ PCT_LESS_BACHELOR_DEGREE, data = seattle)
obese_fit_sum <- summary(obese_fit)
obese_fit_sum

```

```{r, asthma}

asthma <- seattle %>%
ggplot(aes(x = PCT_ENGLISH_LESSTHAN_VERY_WELL, y = PCT_ADULT_WITH_ASTHMA)) +
 geom_point(color = "turquoise4") +
 labs(title = "Lack of English Proficiency versus Asthma",
      x = "Percentage of English Language Learners",
      y = "Percentage of People who have Asthma") +
 theme_minimal() +
 geom_smooth(method = lm, se = FALSE)
asthma
asthma_fit <- lm(PCT_ADULT_WITH_ASTHMA ~ PCT_LESS_BACHELOR_DEGREE, data = seattle)
asthma_fit_sum <- summary(asthma_fit)
asthma_fit_sum

# this one has the smallest r^2 out of the factors I've looked at. Clearly not going to be statistically significant.

```

```{r, socioeconomic}

socioeconomic <- seattle %>%
ggplot(aes(x = PCT_ENGLISH_LESSTHAN_VERY_WELL, y = SOCIOECONOMIC_PERCENTILE)) +
 geom_point(color = "slateblue4") +
 labs(title = "Lack of English Proficiency vs Socioeconomic Disadvantage",
      x = "Percentage of English Language Learners",
      y = "Socioeconomic Disadvantage") +
 theme_minimal() +
 geom_smooth(method = lm, se = FALSE)
socioeconomic
socioeconomic_fit <- lm(SOCIOECONOMIC_PERCENTILE ~ PCT_LESS_BACHELOR_DEGREE, data = seattle)
socioeconomic_fit_sum <- summary(socioeconomic_fit)
socioeconomic_fit_sum

```

```{r, mental}

mental <- seattle %>%
ggplot(aes(x = PCT_ENGLISH_LESSTHAN_VERY_WELL, y = PCT_ADULTMENTALHEALTHNOTGOOD)) +
 geom_point(color = "lightblue4") +
 labs(title = "Lack of English Proficiency versus Poor Mental Health",
      x = "Percentage of English Language Learners",
      y = "Percentage of People with Poor Mental Health") +
 theme_minimal() +
 geom_smooth(method = lm, se = FALSE)
mental
mental_fit <- lm(PCT_ADULTMENTALHEALTHNOTGOOD ~ PCT_LESS_BACHELOR_DEGREE, data = seattle)
mental_fit_sum <- summary(mental_fit)
mental_fit_sum

```

```{r, diabetes}

diabetes <- seattle %>%
ggplot(aes(x = PCT_ENGLISH_LESSTHAN_VERY_WELL, y = PCT_ADULT_WITH_DIABETES)) +
 geom_point(color = "dodgerblue4") +
 labs(title = "Lack of English Proficiency versus Incidence of Diabetes",
      x = "Percentage of English Language Learners",
      y = "Percentage of Adults with Diabetes") +
 theme_minimal() +
 geom_smooth(method = lm, se = FALSE)
diabetes
diabetes_fit <- lm(PCT_ADULTMENTALHEALTHNOTGOOD ~ PCT_ADULT_WITH_DIABETES, data = seattle)
diabetes_fit_sum <- summary(diabetes_fit)
diabetes_fit_sum

```
